---
title: 'Lab 9 - Part 2 : chi-square lizard tails'
author: "Chase Tarr"
date: "11/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)

```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizard species:
```{r}
lizards %>% 
  count(spp) %>% # counting species number
  arrange(-n) #arrange based on largest species
```

## Make a subset
 - UTST lizards
 - Sites: IBPE (enclosure grassland site), RABB (rabbit mesquite site), "EAST" (east tarbush site)
 - only want to keep observations where tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("W", "B"))
```

## Find counts and proportions of lizards with broken or whole tails at each site
 - count tells for each site, how many lizards had a recorded broken or whole tail
```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% #updates percentages to 2 digits
  janitor::adorn_ns() # includes count in each cell for table
```

## Convert to a contingency table
 - Use `column_to_rowname()`
 
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square
```{r}
lizard_x2 <- chisq.test(lizard_ct)

lizard_x2

lizard_tidy <- broom::tidy(lizard_x2) # for in line referencing to easily call values
```
chi-sq test statistic value (13.06): aggregate value indicating how different the values in our table are from the expected null hypothesis scenario values

p-value of 0.0015 = probability that i would get counts that are at least this different from the null hypothesis scenario if they were actually from independent categorical variables. Yes we have sufficient evidence to reject the null hypothesis of independence, meaning we would retain the alternative hypothesis that site and tail breakage are not independent.

Site and Lizard tail breakage are not independent of one another.

in line referencing: 

There is a significant association (i.e. non-independence) between site and lizard tail breakage($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`)
 - can round these, look at key for help
  - remember what you are asking for when you type the final (significant association line)

hello






